---
title: "project_a"
author: "maria"
date: "9/20/2017"
output:
  html_document: default
  pdf_document: default
---
```{r packages, echo = FALSE, message=FALSE, warning=FALSE,}
library(readxl)
library(ggplot2)
library(knitr)
library(dplyr)
library(rlang)
library(dplyr)
library(tidyr)
# BEGIN libraries for map !!! You might need to install these ones
library(rgdal)
library(plyr)
library(rgeos)
# END libraries for map

```

#Analysing crime in London

```{r, echo = FALSE}
crime <- read_excel("Official_data/london-ward-well-being-probability-scores.xls", sheet = "Data")
# data by borough
crime_borough <- crime[c(2,629:660),]
crime_borough[,5:ncol(crime_borough)] <- sapply(crime_borough[,5:ncol(crime_borough)],as.numeric)
# data by ward
crime_ward <- crime[2:626,]
crime_ward[,5:ncol(crime_ward)] <- sapply(crime_ward[,5:ncol(crime_ward)],as.numeric)
summary(crime_ward[,29])


```


# Introduction
Modern policy regarding crime rate has transitioned to a focus on preventative measures as opposed to consequential punishment. As outlined by Crawford and Evans (2017) these policies focus on early intervention and improving well-being to secure community safety. A comprehensive understanding of demographic variables that are highly correlated and plausibly causing crime rates is necessary for developing robust policies. Crime rate in London varies drastically by area. In 2013, crime rates ranged from 13.93 to 222.92, with a mean of 82.40 and a standard deviation of 30. With such drastically different levels of crime and significant outliers, there must be an underlying explanation.

The dataset used in this report is from London datastore, an open data-sharing platform with statistics provided from the London mayoral office. The data is from 2013 and its original purpose was to assign a well-being score to different wards in London through measurements of 12 different indicators such as childhood obesity and deliberate fires. However, this report will focus on crime rates in London and explore the relationships between these rates and other demographic features of wards. It will attempt to model crime based on these additional indicators of well-being. Any significant results may prove to be insightful for policy makers considering local crime rates.

##Description
Information on the 12 key indicators of well-being are given for 625 wards in London over a four-year time frame from 2009 to 2013. Additional borough-wide information is provided for the 32 London boroughs. The variables are split into 8 different categories, detailed below, aiming to give a holistic view of the standard of well-being in the different areas of the capital.

###Variables
#Description
Information on the 12 key indicators of well-being are given for 625 wards in London over a four-year time frame from 2009 to 2013. Additional borough-wide information is provided for the 32 London boroughs. The variables are split into 8 different categories, detailed below, aiming to give a holistic view of the standard of well-being in the different areas of the capital.

#Variables


```{r, echo = FALSE}
Variables <- read_excel("./Official_data/Variables.xlsx")[,1:3]
Variables[2:3,1] <- ""
Variables[2:3,1] <- ""
Variables[6,1] <- ""
kable(Variables)
```


##Strength and limitations


## Overview analysis
### Crime by borough

From 2009 to 2013, the crime rates of all boroughs in London have decreased. As seen in the maps of figure x, the boroughs in the outskirts of London seem to have a lower crime rate than the boroughs in the center of London with exception to the City of London. 


```{r crime_by_borough, message=FALSE, warning=FALSE, results='hide',echo=FALSE,fig.cap="Crime rate by borough from 2009 to 2013"}
#selects relevant columns: borough (4) and all crime levels (25:29) and stores in crime_data_by_borough. Note boroughs are repeated because data is for wards

crime_data_by_borough <- crime_borough[c(4,25:29)]

#adds new column names to crime_data_by_borough
colnames(crime_data_by_borough) <- c("borough", "2009", "2010", "2011","2012","2013")

#calculates the average crime level data per year for each borough
crime_data_borough_averages <- aggregate(crime_data_by_borough[, 2:6], list(crime_data_by_borough$borough), mean)

#adds new column names to crime_data_borough_averages
colnames(crime_data_borough_averages) <- c("borough", "2009", "2010", "2011","2012","2013")

#calculates the growth of crime per borough
growth <- (crime_data_borough_averages$`2013` - crime_data_borough_averages$`2009`)/crime_data_borough_averages$`2009`*100

#Adds growth vector to crime_data_borough_averages
crime_data_borough_averages$growth <- growth

#Reads boroughs data as SpatialPolygonsDataFrame
london_boroughs <- readOGR(dsn = "Official_data/Map_data/LondonBoroughs.shp") 

#joins london_boroughs@data with crime_data_borough_averages based on the borough
london_boroughs@data <- left_join(london_boroughs@data, crime_data_borough_averages, by = c('name' = 'borough'))

london_boroughs_f <- fortify(london_boroughs)
# allocate an id variable to the sp d
london_boroughs$id <- row.names(london_boroughs)
# join the data
london_boroughs_f <- left_join(london_boroughs_f, london_boroughs@data) 

#changes the structure of the dataset to be able to plot it to a graph
london_boroughs_f_long <- gather(london_boroughs_f, crime_rate_year, crime_rate, -long, -lat, -order, -hole,-piece,-id,-group,-ons_label,-name,-Partic_Per,-Pop_2001, -PopDensity, -AREA, -PERIMETER, -PopDen, -growth)

ggplot(data = london_boroughs_f_long, # the input data
       aes(x = long, y = lat, fill = crime_rate, group = group)) + # define variables
  geom_polygon() + # plot the boroughs
  geom_path(colour="white", lwd=0.05) + # borough borders
  coord_equal() + # fixed x and y scales
  facet_wrap(~ crime_rate_year) + # one plot per year
  scale_fill_gradient2(low = "#64B5F6",high = "#2196F3", # colors
                        name = "Crime rate") + # legend options
  theme(axis.text = element_blank(), # change the theme options
        axis.title = element_blank(), # remove axis titles
        axis.ticks = element_blank()) # remove axis ticks



```

City of London, remains with the lowest rate of crime across all boroughs with only 14 offences per 1000 daytime inhabitants in 2013. In contrast, its neighbouring boroughs as Lambeth, Hackney, Southwark and Islington continue to present a crime rate of 100. The map below shows the quartile with the highest number of crime rates in London (crime rate over 90). These boroughs are mainly located to the upper and lower side from the City of London. 

```{r boroughs_with_high_crime_rate, message=FALSE, warning=FALSE, results='hide',echo=FALSE, fig.height = 4, fig.width = 5, fig.cap="Crime rate > 90 (2013)"}
#plots the boroughs with highest crime rate of 100
ggplot(london_boroughs_f, aes(long, lat, group = group, fill=`2013`>90)) +
  geom_polygon() + geom_path(colour="white", lwd=0.05) + coord_equal() +
  labs(x = "lat", y = "lon",
       fill = "Crime rate > 90") +
  scale_fill_manual(values=c("#BDBDBD","#2196F3")) +
  ggtitle("Crime rate > 90 (2013) ") +
  theme(axis.text = element_blank(), # change the theme options
        axis.title = element_blank(), # remove axis titles
        axis.ticks = element_blank()) # remove axis ticks
```

On average, boroughs have lowered their crime rates to 19% from 2009 to 2013. The boroughs that seem to be showing the highest decrease in crime rates are on the further southeast of London, as for instance in Greenwich, Barking and Dagenham and Lewisham. The boroughs with the least progress in decreasing crime are Tower Hamlets, Enfield and Kensington and Chelsea. 


```{r crime_rate_growth, message=FALSE, warning=FALSE, results='hide',echo=FALSE, fig.height = 4, fig.width = 5, fig.cap="Crime rate growth from 2009 to 2013"}
ggplot(data = london_boroughs_f, # the input data
       aes(x = long, y = lat, fill = growth, group = group)) + # define variables
  geom_polygon() + # plot the boroughs
  geom_path(colour="white", lwd=0.05) + # borough borders
  ggtitle("Crime rate growth from 2009 to 2013 ") +
  coord_equal() + # fixed x and y scales
  scale_fill_gradient2(low = "#66BB6A",high = "#4CAF50", # colors
                        name = "Growth (%)") + # legend options
  theme(axis.text = element_blank(), # change the theme options
        axis.title = element_blank(), # remove axis titles
        axis.ticks = element_blank()) # remove axis ticks
```

### Crime historically

###AMR###

###TEXT: added some intro text. Check with Maria if it is needed for report###
Crime, in general, is very hard to measure and as in various big cities around the world, measuring crime in London is no easy task. Police statistics is usually the only measure to local crime. However, police crime figures are not always accurate as they tend to understate recorded crime statistics. Accordingly, crime statistics needs to be investigated with caution; sometimes data from hospitals intakes or victim surveys infer very different theories. Geographically, London is divided into 32 boroughs and each borough consists of multiple wards. During our analysis, we will look at crimes rates across the 32 boroughs along 5 years; from 2009 till 2013.

Analyzing crime data across London benefit both the governmental organizations and the individuals. For example, organizations can use the data to implement better services and check for core reasons why crime in one area is higher than the other, while, individuals can utilize this data to make calculated decisions on where to buy or rent a house in London. 

Our dataset measures the well-being scores across all London's wards. In our analysis, we will look at how different factors affect crime rates in different areas, such as how do unemployment rate affect crime rates or how GSCE test score relate to the crime rates. We will investigate the relationship between crime and different factors using linear regression tests and other statistical tests such as hypothesis test.
###TEXT###

```{r}

# selects relevant data from the whole dataset. Borough and crime rate on each of the 5 years.
crime_by_year <- crime_borough[, c(4, 25:29)]

# changes the names of the coloumns
colnames(crime_by_year) <- c("Borough", "Crime_Rate_2009", "Crime_Rate_2010", "Crime_Rate_2011","Crime_Rate_2012","Crime_Rate_2013")

# plot the crime rate of 2009 for each borough
ggplot(crime_by_year, aes(Borough, Crime_Rate_2009)) +
  geom_bar(stat = "identity") +
  theme(axis.text = element_text(angle = 45, hjust = 1)) +
  ylim(0, 150)

# plot the crime rate of 2010 for each borough
ggplot(crime_by_year, aes(Borough, Crime_Rate_2010)) +
  geom_bar(stat = "identity") +
  theme(axis.text = element_text(angle = 45, hjust = 1)) +
  ylim(0, 150)

# plot the crime rate of 2011 for each borough
ggplot(crime_by_year, aes(Borough, Crime_Rate_2011)) +
  geom_bar(stat = "identity") +
  theme(axis.text = element_text(angle = 45, hjust = 1)) +
  ylim(0, 150)

# plot the crime rate of 2012 for each borough
ggplot(crime_by_year, aes(Borough, Crime_Rate_2012)) +
  geom_bar(stat = "identity") +
  theme(axis.text = element_text(angle = 45, hjust = 1)) +
  ylim(0, 150)

# plot the crime rate of 2013 for each borough
ggplot(crime_by_year, aes(Borough, Crime_Rate_2013)) +
  geom_bar(stat = "identity") +
  theme(axis.text = element_text(angle = 45, hjust = 1)) +
  ylim(0, 150)

###Ask Valeria on how to put the 5 plots on one plot (just like the maps one)###

summary(crime_by_year)


#Average crime over the 5 years, per borough
average <- crime_by_year %>% mutate(average = (Crime_Rate_2009 + Crime_Rate_2010 + Crime_Rate_2011 + Crime_Rate_2012 + Crime_Rate_2013)/ 5)


ggplot(average, aes(Borough, average)) +
  geom_bar(stat = "identity") +
  theme(axis.text = element_text(angle = 45, hjust = 1)) +
  ylim(0, 150)

ggplot(average, aes(Borough, average)) +
  geom_point() +
  theme(axis.text = element_text(angle = 45, hjust = 1))

```

Looking at crime rates across the five years, we can see clear trend of crime rates decreasing from 2009 till 2013 across all the London boroughs. The bar plots, for each year, show the crim rates per borough for that specific year. If you compare the 5 plots, you will notice that the crime rates are decreasing year by year.

The scatter plot shows the average crime rate (over the 5 years) across each borough. One thing to note, is the City of London, which has the lowest crime rate across all of London. 


#Correlation Analysis
The correlation analysis in this report adopts the Pearson correlation coefficient with the aim to examine the strength and direction of the linear relationship between two variables, i.e. crime rate and one other factors.The correlation coefficient ranges from −1 to +1, and stronger relationship between two variables as larger absolute value of the coefficient. Positive coefficient means both variables tend to increase or decrease together whereas negetive coefficient means two variables have opposite directions (Zhou et al., 2016).

The following correlation analysis does not conclude that changes in one variable cause changes in crime rate or changes in creime rate cause in other variable. This analysis will not tell whether a relationship is causal.

```{r}
#the following loop can produce long table 
var_list = c("life_expectancy","childhood_obesity","incapacity_benefit","unemployment_rate","crime_rate","deliberate_fires","GCSE_point","unauthorised_absence","dependent_children","public_transport","home_access")

crime_ward_long = data.frame()
for (i in 1:length(var_list)) {
  temp <- select(crime_ward, Ward, Borough, (i*5):(i*5+4))
  colnames(temp)[3:7] <- c(2009:2013)
  tidytemp <- gather_(temp, "year", var_list[i], colnames(temp)[3:7])
  if (i==1) {
    crime_ward_long <- tidytemp
  } else {
    crime_ward_long <- cbind(crime_ward_long, tidytemp[,4])
  }
}
gb <- crime_ward_long %>% dplyr::group_by(year)

```
```{r}
# the codes below define a function that allows you to make scatters boxplot with correlation on them
# argument: col_name, string with double quotes, e.g. "unemployment_rate"
cor_plot <- function(col_name) {
  corr <- dplyr::summarise(gb,r=round(cor(!!sym(col_name),crime_rate),4))
  print(ggplot(crime_ward_long,aes_string(x=col_name,y="crime_rate"))+geom_point(alpha=0.2)+
    facet_wrap(~ year, nrow=2) +
    geom_smooth() +
    geom_text(data=corr, aes(label=paste("r=",r,sep="")),x=quantile(gb[[col_name]],0.9),y=270) )
}
# cor_plot("unemployment_rate")
```
##Correlations between crime rate and 3 health-related indexes
###Description
These three health-realated indexes including rolling 5-year combined life expectancies, prevalence of obsity by area of child residence, and incapcaity benefit or servere disablement allowance claimant rate.
###Life Expectancy
According to the data description provided by London datastore, rolling 5-year combined life expectancies instead of simple one-year life expectancies are used for wards to reduce the efffects of the variability in number of deaths in each year.



```{r}
cor_plot("life_expectancy")
ggplot(gb, aes(x=cut(childhood_obesity, breaks =10), y=crime_rate))+geom_boxplot()+facet_wrap(~year)
```
From the above scatter points in different years, moderate negative relationship between crime rate and life expectancy can be found, with Pearson r from -0.3631 to 0.4545 in 5 different years. Also, a significant portion of points are far from the smooth line indicates only a moderate linear relationship between the variables.

###Childhood Obesity
This estimates use the latest three years of National Child Measurement Programme (NCMP) data combined. This means, the data in 2013 is the average of data in 2011, 2012, and 2013. In addition, children with a Body Mass Index (BMI) greater than or equal to the 95th centile of the British 1990 growth reference (UK90) BMI distribution have been classified as obese. 
```{r}
cor_plot("childhood_obesity")
ggplot(gb, aes(x=cut(childhood_obesity, breaks =5), y=crime_rate))+geom_boxplot()+facet_wrap(~year)
```
Similar to the correlation between crime rate and life expectancy, crime rate and childhood obesity share a moderate positive relationship, with Pearson r's range from 0.372 to 0.389.
It is noticebaly that expect for in year 2013, wards that have extremely high childhood obesity score tend to have lower crime rate, the phenomenon violate the moderate positive relationship between the two variables.

###Incapacity Benefit
Incapacity Benefit rate is the rate of paying Incapacity Benefit (IB) to those who aged 16-64 and unable to work for 28 weeks in a row or more because of illness or disability by Severe Disablement Allowance (SDA).
```{r}
cor_plot("incapacity_benefit")
ggplot(gb, aes(x=cut(childhood_obesity, breaks =5), y=crime_rate))+geom_boxplot()+facet_wrap(~year)
```
Again crime rate and incapacity benefit rate have a moderate positive relationship with Pearson r's range from 0.3917 to 0.4752 in different years.

In conclusion, moderate correlations are found between crime rate and three health-related rates. The moderate correlations may be explained by many extreme data values since the Pearson correlation coefficient is very sensitive to extreme data values. A single value that is very different from the other values in a data set can greatly change the value of the coefficient. Many outliers, shown on the the above boxplots, may greatly affect the result of the correlation between crime rate and other variable.

##Linear regression

##Conclusion and analysis









