---
title: "project_a"
author: "maria"
date: "9/20/2017"
output:
  pdf_document: default
  html_document: default
---
```{r packages, echo = FALSE, message=FALSE, warning=FALSE,}
library(readxl)
library(ggplot2)
library(knitr)
library(dplyr)
library(rlang)
library(dplyr)
library(ggplot2)
library(tidyr)
# BEGIN libraries for map !!! You might need to install these ones
library(rgdal)
library(plyr)
library(rgeos)
# END libraries for map

```

#Examining the relationship between crime rates and well-being in London

```{r, echo = FALSE}
crime <- read_excel("Official_data/london-ward-well-being-probability-scores.xls", sheet = "Data")
# data by borough
crime_borough <- crime[c(2,629:660),]
crime_borough[,5:ncol(crime_borough)] <- sapply(crime_borough[,5:ncol(crime_borough)],as.numeric)
# data by ward
crime_ward <- crime[2:626,]
crime_ward[,5:ncol(crime_ward)] <- sapply(crime_ward[,5:ncol(crime_ward)],as.numeric)
summary(crime_ward[,29])


```


## Introduction
Modern policy regarding crime rate has transitioned to a focus on preventative measures as opposed to consequential punishment. As outlined by Crawford and Evans (2017) these policies focus on early intervention and improving well-being to secure community safety. A comprehensive understanding of demographic variables that are highly correlated and plausibly causing crime rates is necessary for developing robust policies. Crime rate in London varies drastically by area. In 2013, crime rates ranged from 13.93 to 222.92, with a mean of 82.40 and a standard deviation of 30. With such drastically different levels of crime and significant outliers, there must be an underlying explanation.

The dataset used in this report is from London datastore, an open data-sharing platform with statistics provided from the London mayoral office. The data is from 2013 and its original purpose was to assign a well-being score to different wards in London through measurements of 12 different indicators such as childhood obesity and deliberate fires. However, this report will focus on crime rates in London and explore the relationships between these rates and other demographic features of wards. It will attempt to model crime based on these additional indicators of well-being. Any significant results may prove to be insightful for policy makers considering local crime rates.

<<<<<<< HEAD
##Description
Information on the 12 key indicators of well-being are given for 625 wards in London over a four-year time frame from 2009 to 2013. Additional borough-wide information is provided for the 32 London boroughs. The variables are split into 8 different categories, detailed below, aiming to give a holistic view of the standard of well-being in the different areas of the capital.

###Variables
#Description
Information on the 12 key indicators of well-being are given for 625 wards in London over a four-year time frame from 2009 to 2013. Additional borough-wide information is provided for the 32 London boroughs. The variables are split into 8 different categories, detailed below, aiming to give a holistic view of the standard of well-being in the different areas of the capital.

#Variables

=======
## Dataset description
Information on the 12 key indicators of well-being are given for 625 wards in London over a four-year time frame from 2009 to 2013. Additional borough-wide information is provided for the 32 London boroughs. The variables are split into 8 different categories, detailed below, aiming to give a holistic view of the standard of well-being in the different areas of the capital.

### Variables
>>>>>>> 7a74c1211391c752b51bb834c995e12e0141a878

```{r, echo = FALSE}
Variables <- read_excel("./Official_data/Variables.xlsx")[,1:3]
Variables[2:3,1] <- ""
Variables[2:3,1] <- ""
Variables[6,1] <- ""
kable(Variables)
```


##Strength and limitations

Some text needs to be added here

## Overview analysis
### Crime by borough

From 2009 to 2013, the crime rates of all boroughs in London have decreased. As seen in the maps of figure below, the boroughs in the outskirts of London seem to have a lower crime rate than the boroughs in the center of London with exception to the City of London. 

```{r crime_by_borough, message=FALSE, warning=FALSE, results='hide',echo=FALSE,fig.cap="Crime rate by borough from 2009 to 2013"}
#selects relevant columns: borough (4) and all crime levels (25:29) and stores in crime_data_by_borough. Note boroughs are repeated because data is for wards

crime_data_by_borough <- crime_borough[c(4,25:29)]

#adds new column names to crime_data_by_borough
colnames(crime_data_by_borough) <- c("borough", "2009", "2010", "2011","2012","2013")

#calculates the average crime level data per year for each borough
crime_data_borough_averages <- aggregate(crime_data_by_borough[, 2:6], list(crime_data_by_borough$borough), mean)

#adds new column names to crime_data_borough_averages
colnames(crime_data_borough_averages) <- c("borough", "2009", "2010", "2011","2012","2013")

#calculates the growth of crime per borough
growth <- (crime_data_borough_averages$`2013` - crime_data_borough_averages$`2009`)/crime_data_borough_averages$`2009`*100

#Adds growth vector to crime_data_borough_averages
crime_data_borough_averages$growth <- growth

#Reads boroughs data as SpatialPolygonsDataFrame
london_boroughs <- readOGR(dsn = "Official_data/Map_data/LondonBoroughs.shp") 

#joins london_boroughs@data with crime_data_borough_averages based on the borough
london_boroughs@data <- left_join(london_boroughs@data, crime_data_borough_averages, by = c('name' = 'borough'))

london_boroughs_f <- fortify(london_boroughs)
# allocate an id variable to the sp d
london_boroughs$id <- row.names(london_boroughs)
# join the data
london_boroughs_f <- left_join(london_boroughs_f, london_boroughs@data) 

#changes the structure of the dataset to be able to plot it to a graph
london_boroughs_f_long <- gather(london_boroughs_f, crime_rate_year, crime_rate, -long, -lat, -order, -hole,-piece,-id,-group,-ons_label,-name,-Partic_Per,-Pop_2001, -PopDensity, -AREA, -PERIMETER, -PopDen, -growth)

ggplot(data = london_boroughs_f_long, # the input data
       aes(x = long, y = lat, fill = crime_rate, group = group)) + # define variables
  geom_polygon() + # plot the boroughs
  geom_path(colour="white", lwd=0.05) + # borough borders
  coord_equal() + # fixed x and y scales
  facet_wrap(~ crime_rate_year) + # one plot per year
  scale_fill_gradient2(low = "#64B5F6",high = "#2196F3", # colors
                        name = "Crime rate") + # legend options
  theme(axis.text = element_blank(), # change the theme options
        axis.title = element_blank(), # remove axis titles
        axis.ticks = element_blank()) # remove axis ticks



```

City of London, remains with the lowest rate of crime across all boroughs with only 14 offences per 1,000 daytime inhabitants in 2013. In contrast, its neighbouring boroughs as Lambeth, Hackney, Southwark and Islington continue to present a crime rate of over 100. The map below shows the quartile with the highest number of crime rates in London (crime rate over 90). These boroughs are mainly located to the upper and lower side from the City of London. 

```{r boroughs_with_high_crime_rate, message=FALSE, warning=FALSE, results='hide',echo=FALSE, fig.height = 4, fig.width = 5, fig.cap="Crime rate > 90 (2013)"}
#plots the boroughs with highest crime rate of 100
ggplot(london_boroughs_f, aes(long, lat, group = group, fill=`2013`>90)) +
  geom_polygon() + geom_path(colour="white", lwd=0.05) + coord_equal() +
  labs(x = "lat", y = "lon",
       fill = "Crime rate > 90") +
  scale_fill_manual(values=c("#BDBDBD","#2196F3")) +
  ggtitle("Crime rate > 90 (2013) ") +
  theme(axis.text = element_blank(), # change the theme options
        axis.title = element_blank(), # remove axis titles
        axis.ticks = element_blank()) # remove axis ticks
```

On average, boroughs have lowered their crime rates by 19% from 2009 to 2013. The boroughs that seem to be showing the highest decrease in crime rate are on the further southeast of London, as for instance in Greenwich, Barking and Dagenham and Lewisham. The boroughs with the least progress in decreasing crime are Tower Hamlets, Enfield and Kensington and Chelsea. 

```{r crime_rate_growth, message=FALSE, warning=FALSE, results='hide',echo=FALSE, fig.height = 4, fig.width = 5, fig.cap="Decrease of crime rates by borough from 2009 to 2013"}
ggplot(data = london_boroughs_f, # the input data
       aes(x = long, y = lat, fill = growth, group = group)) + # define variables
  geom_polygon() + # plot the boroughs
  geom_path(colour="white", lwd=0.05) + # borough borders
  ggtitle("Decrease of crime rates by borough from 2009 to 2013 ") +
  coord_equal() + # fixed x and y scales
  scale_fill_gradient2(low = "#66BB6A",high = "#4CAF50", # colors
                        name = "Decrease (%)") + # legend options
  theme(axis.text = element_blank(), # change the theme options
        axis.title = element_blank(), # remove axis titles
        axis.ticks = element_blank()) # remove axis ticks
```

### Crime historically

##Theory

##Correlations
###Correlations with crime rate

#Theory

#Correlations
```{r}
<<<<<<< HEAD
#the following codes can produce long table and plots with correlations on them
# correlation part not finished
var_list = c("childhood_obesity","incapacity_benefit","unemployment_rate","crime_rate","deliberate_fires","GCSE_point","unauthorised_absence","dependent_children","public_transport","home_access")

crime_ward_long = data.frame()
for (i in 1:length(var_list)) {
  temp <- select(crime_ward, Ward, Borough, (i*5+5):(i*5+9))
  colnames(temp)[3:7] <- c(2009:2013)
  tidytemp <- gather_(temp, "year", var_list[i], colnames(temp)[3:7])
  if (i==1) {
    crime_ward_long <- tidytemp
  } else {
    crime_ward_long <- cbind(crime_ward_long, tidytemp[,4])
  }
}
gb <- crime_ward_long %>% dplyr::group_by(year)
for (item in var_list) {
  if (item!="crime_rate") {
    corr <- dplyr::summarise(gb,r=round(cor(!!sym(item),crime_rate),4))
    print(ggplot(crime_ward_long,aes_string(x=item,y="crime_rate"))+geom_point(alpha=0.5)+
      facet_wrap(~ year, nrow=2) +
      geom_smooth() +
      geom_text(data=corr, aes(label=paste("r=",r,sep="")),x=quantile(gb[[item]],0.9),y=270) )
  }
}

=======

# selects relevant data from the whole dataset. Borough and crime rate on each of the 5 years.
crime_by_year <- crime_borough[, c(4, 25:29)]

# changes the names of the coloumns
colnames(crime_by_year) <- c("Borough", "Crime_Rate_2009", "Crime_Rate_2010", "Crime_Rate_2011","Crime_Rate_2012","Crime_Rate_2013")

# plot the crime rate of 2009 for each borough
ggplot(crime_by_year, aes(Borough, Crime_Rate_2009)) +
  geom_bar(stat = "identity") +
  theme(axis.text = element_text(angle = 45, hjust = 1)) +
  ylim(0, 150)

# plot the crime rate of 2010 for each borough
ggplot(crime_by_year, aes(Borough, Crime_Rate_2010)) +
  geom_bar(stat = "identity") +
  theme(axis.text = element_text(angle = 45, hjust = 1)) +
  ylim(0, 150)

# plot the crime rate of 2011 for each borough
ggplot(crime_by_year, aes(Borough, Crime_Rate_2011)) +
  geom_bar(stat = "identity") +
  theme(axis.text = element_text(angle = 45, hjust = 1)) +
  ylim(0, 150)

# plot the crime rate of 2012 for each borough
ggplot(crime_by_year, aes(Borough, Crime_Rate_2012)) +
  geom_bar(stat = "identity") +
  theme(axis.text = element_text(angle = 45, hjust = 1)) +
  ylim(0, 150)

# plot the crime rate of 2013 for each borough
ggplot(crime_by_year, aes(Borough, Crime_Rate_2013)) +
  geom_bar(stat = "identity") +
  theme(axis.text = element_text(angle = 45, hjust = 1)) +
  ylim(0, 150)

###Ask Valeria on how to put the 5 plots on one plot (just like the maps one)###

summary(crime_by_year)


#Average crime over the 5 years, per borough
average <- crime_by_year %>% mutate(average = (Crime_Rate_2009 + Crime_Rate_2010 + Crime_Rate_2011 + Crime_Rate_2012 + Crime_Rate_2013)/ 5)


ggplot(average, aes(Borough, average)) +
  geom_bar(stat = "identity") +
  theme(axis.text = element_text(angle = 45, hjust = 1)) +
  ylim(0, 150)

ggplot(average, aes(Borough, average)) +
  geom_point() +
  theme(axis.text = element_text(angle = 45, hjust = 1))

```

Looking at crime rates across the five years, we can see clear trend of crime rates decreasing from 2009 till 2013 across all the London boroughs. The bar plots, for each year, show the crim rates per borough for that specific year. If you compare the 5 plots, you will notice that the crime rates are decreasing year by year.

The scatter plot shows the average crime rate (over the 5 years) across each borough. One thing to note, is the City of London, which has the lowest crime rate across all of London. 


# Correlation Analysis
The correlation analysis in this report adopts the Pearson correlation coefficient, denoted r, with the aim to examine the strength and direction of the linear relationship between two continuous variables, i.e. crime rate and one of the other rates.The correlation coefficient ranges from −1 to +1, and a stronger relationship between two variables as a larger absolute value of the coefficient. Positive coefficient means both variables tend to increase or decrease together whereas negetive coefficient means two variables have opposite directions (Zhou et al., 2016).

The following correlation analysis does not conclude that changes in one variable cause changes in crime rate or changes in creime rate cause in other variable. This analysis will not tell whether a relationship is causal;however, these will be answered in the regression part.


##Test of correlation 
The following part will test the significant of the correlation between crime rates and other variables.

```{r}
cor_test_pvalue <- c()
cor_test_cor <- c()
variable_name<-c("life_expectancy","childhood_obesity","incapacity_benefit","unemployment_rate","deliberate_fires","GCSE_point","unauthorised_absence","dependent_children","public_transport","home_access", "subjective_well_being")
for (i in seq(from = 9, to = 64, by= 5 )){
  if (i != 29){
  cor_test <- cor.test(crime_ward[[i]],crime_ward[[29]])
  cor_test_pvalue <-c(cor_test_pvalue,(cor_test$p.value))
  cor_test_cor <- c(cor_test_cor, (cor_test$estimate[[1]]))
  }
}

cor_test_df <- data.frame(variable_name,cor_test_cor,cor_test_pvalue)
cor_test_df <- cor_test_df %>% arrange(desc(cor_test_pvalue))
names(cor_test_df)[1]<-paste("Variable Name")
names(cor_test_df)[2]<-paste("Correlation Coefficient")
names(cor_test_df)[3]<-paste("P-Value")
kable(cor_test_df)
```

We would like to examine whether the correlation between crime rate and one of the other variables is significant by comparing the p-value to a certain significance level, denoted α. 

The significance level refers to a certain degree of risk of concluding that a diffrence exists while in fact there is no diffrence. We belive a significance level of 0.05 is fair enough for this report. Therefore, if the p-value of a correlation test is less than or equal to α, we would conclude that the correlation is statistically significant and would be further discussed.
We then remove subjective well-being score and the proportion of accessing to public open space and nature from our further discussions. In addition, we exclude deliberate fire becasue it has a relatively low correlation with crime rate (approximatedly 0.13), the rate which is not significant enough as a rule of thumb.

Therefore, we would only focus on the following variables in the following analysis.

```{r}
# crime_ward_2013
crime_ward_2013 <- crime_ward[,c(3,4,seq(9,64,5))]
colnames(crime_ward_2013) <- c("Ward", "Borough","life_expectancy","childhood_obesity","incapacity_benefit","unemployment_rate","crime_rate","deliberate_fires","GCSE_point","unauthorised_absence","dependent_children","public_transport","home_access","well-being")
# the codes below define a function that allows you to make scatters boxplot with correlation on them
# argument: col_name, string with double quotes, e.g. "unemployment_rate"
cor_plot <- function(col_name) {
  corr <- data.frame(r=round(cor(crime_ward_2013[[col_name]],crime_ward_2013[,"crime_rate"]),4))
  print(ggplot(crime_ward_2013,aes_string(x=col_name,y="crime_rate")) + geom_point(alpha=0.2) +
    geom_smooth(method = "lm") +
    geom_text(data=corr, aes(label=paste("r=",crime_rate,sep="")),x=quantile(crime_ward_2013[[col_name]],0.9),y=200) )

}
# example: cor_plot("unemployment_rate")
```

### Life Expectancy

According to the data description provided by London datastore, rolling 5-year combined life expectancies instead of simple one-year life expectancies are used for wards to reduce the efffects of the variability in number of deaths in each year.

A study in 2015 reviewed that biological aging processes may be impacted by socioeconomic milieu, meaning that those who living in noisy area with high crime may have shorter cellular aging estimated by leukocyte telomere length (Park et al., 2015). Therefore, it is reasonable to estimate that a negative correlation between crime rate and life expectancy in 2013.


```{r}
cor_plot("life_expectancy")
```

From the above scatter point graph, moderate negative relationship between crime rate and life expectancy can be found, with Pearson r -0.3631. Also, a significant portion of points are far from the smooth line indicates only a moderate linear relationship between the variables.

### Childhood Obesity

Children with a Body Mass Index (BMI) greater than or equal to the 95th centile of the British 1990 growth reference (UK90) BMI distribution have been classified as obese. The Childhood Obesity rate are combined by the latest three years of National Child Measurement Programme (NCMP) data. This means, the data we use is the average of data in 2011, 2012, and 2013. 

An et al. (2017) revealed a  modest relationship between residential neighbourhood safety and weight-related behaviours/outcomes among children and adolescents through examing 22 acdemic studies, half of which were conducted in the USA, and one of which was conducted in a developing country. There analysis suggested that living in district with relatively high crime rate was associated with a reduction in children's physical activity, and therefore gain in body mass index, but no change in childhood obesity risk.

Therefore, we expect childhood obesity rate and crime rate have moderate positive correlation.

```{r}
cor_plot("childhood_obesity")
```

Similar to the correlation between crime rate and life expectancy, crime rate and childhood obesity share a moderate positive relationship, with r of 0.3757.

The analysis results confirm our expectation based on the study conducted by An et al. (2017). It is noticebaly that expect for in year 2013, wards that have extremely high childhood obesity score tend to have lower crime rate, the phenomenon violate the moderate positive relationship between the two variables.

###Incapacity Benefit

Incapacity Benefit rate represents Incapacity Benefit or Severe Disablement Allowance claimant rate. Incapacity Benefit (IB) is paid to people who are incapable of work and who meet certain contribution conditions and Severe Disablement Allowance (SDA) is paid to those unable to work for 28 weeks in a row or more because of illness or disability.

```{r}
cor_plot("incapacity_benefit")
```

Again crime rate and incapacity benefit rate have a moderate positive relationship, with a r equals to 0.3917 in 2013. The moderate correlation may be explained by many extreme data values since the Pearson correlation coefficient is very sensitive to extreme data values. A single value that is very different from the other values in a data set can greatly change the value of the coefficient. 


###Unemployment Rate
Unemployment rate here refers to percentage of working-age residents claiming Jobseeker's Allowance (JSA) or National Insurance Credits. JSA is a benefit payable to unemployed people so presumably the percentage can be used to estimate unemployment rate. 
A study found significantly positive effects of unemployment on property crime rates using US state data (Raphael and Winter‐Ebmer, 2001). Although we do not have crime rate data of different types of crimes (i.e. property crime), we expect the relation of crime rate in general and unemployment rate to be positively correlated.
>>>>>>> e7bc4b8e6c59f486b07b426bd4ba7595a7259d73

"Incapacity Benefit or Severe Disablement Allowance claimant rate. 
Incapacity Benefit (IB) is paid to people who are incapable of work and who meet certain contribution conditions. Severe Disablement Allowance (SDA) is paid to those unable to work for 28 weeks in a row or more because of illness or disability. SDA was removed for new claims in April 2001. Time period used is a snapshot of May from each year. Demominator is population aged 16-64. 
Source: IB/SDA from DWP, Population from GLA projections."
###basic data selecting
```{r}
<<<<<<< HEAD
#read all the names of columns
crime_health <- select(crime_ward, Ward:`Incapacity Benefit rate - 2013`, -(`Life Expectancy 2005-2009`:`Life Expectancy 2008-12`),`Crime rate - 2009`:`Crime rate - 2013`)
crime_health[,3:18] <- sapply(crime_health[,3:18], as.numeric)
crime_health <- mutate(crime_health, `Childhood Obesity 2009-2013`=(`Childhood Obesity 2009`+`Childhood Obesity 2010`+`Childhood Obesity 2011`+`Childhood Obesity 2012`+`Childhood Obesity 2013`)/5)
crime_health <- mutate(crime_health, `Incapacity Benefit rate - 2009-2013`=(`Incapacity Benefit rate - 2009`+`Incapacity Benefit rate - 2010`+`Incapacity Benefit rate - 2011`+`Incapacity Benefit rate - 2012`+`Incapacity Benefit rate - 2013`)/5)
crime_health <- mutate(crime_health, `Crime rate - 2009-2013`=(`Crime rate - 2009`+`Crime rate - 2010`+`Crime rate - 2011`+`Crime rate - 2012`+`Crime rate - 2013`)/5)

crime_health1 <- select(crime_health, -(`Childhood Obesity 2009`:`Crime rate - 2013`))
=======
cor_plot("unemployment_rate")
```
<<<<<<< HEAD
We also expect areas with considerately dense population (i.e. inner London) may potentially have more crimes even if the general employment rate are similar to suburban areas. If we split boroughs into inner and outer London, we can see employment and crime rate show only weak correlations of around 0.2 for innter London but stronger correlations (0.58-0.62) for outer London boroughs. 
```{r}
# split borough into inner and outer
desig <- read.csv("./Official_data/borough_desig.csv")
crime_ward_2013_desig <- merge(crime_ward_2013[2:nrow(crime_ward_2013),],desig,by="Borough")
inner_ward <- crime_ward_2013_desig%>% filter(Designation=="Inner")
outer_ward <- crime_ward_2013_desig%>% filter(Designation=="Outer")
corr <- data.frame(r=round(cor(inner_ward[,"unemployment_rate"],inner_ward[,"crime_rate"]),4))
ggplot(inner_ward,aes(x=unemployment_rate,y=crime_rate))+geom_point(alpha=0.2)+
  geom_smooth(method="lm") +
  geom_text(data=corr, aes(label=paste("r=",r,sep="")),x=quantile(inner_ward$unemployment_rate,0.9),y=200) 

corr <- data.frame(r=round(cor(outer_ward[,"unemployment_rate"],outer_ward[,"crime_rate"]),4))
ggplot(outer_ward,aes(x=unemployment_rate,y=crime_rate))+geom_point(alpha=0.2)+
  geom_smooth(method="lm") +
  geom_text(data=corr, aes(label=paste("r=",r,sep="")),x=quantile(outer_ward$unemployment_rate,0.9),y=200)

mean_desig <- crime_ward_2013_desig%>% group_by(Designation)  %>% dplyr::summarise(avg_crime=mean(crime_rate))
mean_desig
>>>>>>> 252852c94f5c36b32d05520868a9d73c414d2d82
```
=======
Our scatter plot shows moderate positive correlation between the two variables, which confirmed our hypothesis. The wards with higher unemployment rates are likely to have higher crime rates. 

###Education (GCSE Score and Unauthorised Pupil Absence)
GCSE average point scores refers to GCSE and Equivalent point scores for pupils at the end of Key Stage 4 (KS4) in maintained schools. Unauthorised Absence is absence without permission from a teacher or other authorised representative of the school. Wards with higher GCSE scores and lower unauthorised pupil absence suggect higher education well-being. Not surprisingly, these two variables have negatively relationship with a correlation of -0.5979. 
>>>>>>> e7bc4b8e6c59f486b07b426bd4ba7595a7259d73

###Correlation Analysis
```{r}
<<<<<<< HEAD
names(crime_health1)
crime_health2 <- crime_health1[, c(6,3,4,5)]
cor(crime_health2)
=======
# correlation between GCSE & unauthorised pupil absence
cor(crime_ward_2013$GCSE_point, crime_ward_2013$unauthorised_absence)
```
The research of Lochner and Moretti (2004) shows that schooling significantly reduces criminal activity. We expect to see a negative relationship between GCSE score and crime rate but a positive relationship between unauthorised pupil absence. 
```{r}
# plot education varialbes
cor_plot("GCSE_point")
cor_plot("unauthorised_absence")
```
As we can see from the scatters, the relationships are consistant from our expectation. However, correlations (-0.33 and 0.374) between education variables and crime rate are relatively weaker compared with unemployment. Both of the variables are all related to primary school education. If we can also combine data from secondary school or even university to create a complete picture of education, it may reveal a stronger relationship with crime rate. 

### Dependent children
>>>>>>> 252852c94f5c36b32d05520868a9d73c414d2d82

```
###Plots
```{r}
ggplot(crime_health2, aes(x=`Life Expectancy 2009-13`, y=`Crime rate - 2009-2013`))+geom_point()+geom_smooth()
ggplot(crime_health2, aes(x=`Childhood Obesity 2009-2013`, y=`Crime rate - 2009-2013`))+geom_point()+geom_smooth()
ggplot(crime_health2, aes(x=`Incapacity Benefit rate - 2009-2013`, y=`Crime rate - 2009-2013`))+geom_point()+geom_smooth()

ggplot(crime_health2, aes(x=cut(`Life Expectancy 2009-13`, breaks =5), y=`Crime rate - 2009-2013`))+geom_boxplot()
ggplot(crime_health2, aes(x=cut(`Childhood Obesity 2009-2013`, breaks =5), y=`Crime rate - 2009-2013`))+geom_boxplot()
ggplot(crime_health2, aes(x=cut(`Incapacity Benefit rate - 2009-2013`, breaks =5), y=`Crime rate - 2009-2013`))+geom_boxplot()

```
  
  
  
  
  

  

<<<<<<< HEAD
##Linear regression
=======








##Correlation Matrix
Finally, we study the correlation between any two variables by construsting a correlation matrix. Before we get into , we notice that some of our independent variables are highly correlated, suggesting multicollinearity exists which can impact the model fits. 

```{r}
cor_mat_df<-round(cor(select(crime_ward_2013, -Ward, -Borough, -home_access, -`well-being`, -deliberate_fires)),4)
kable(cor_mat_df)

```




	
## Linear regression
>>>>>>> 252852c94f5c36b32d05520868a9d73c414d2d82

##Conclusion and analysis









